---
title: "DATA 607 Project 1"
author: "Albert Gilharry"
date: "February 22, 2018"
output:
  pdf_document:
    toc: yes
  html_document:
    css: ./css.css
    highlight: pygments
    pdf_document: default
    theme: spacelab
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r load-libraries}
library("stringr")
```
## Load Input File
```{r load-data}
lines <- scan("data/tournamentinfo.txt", sep="\n", what="raw")
# preview data
print(lines[1:15])
```

## Initialize Data Structure

<div id = "solution">
I decided to use use a data frame to store the data and leverage its capabilities to do subsetting, calculations, and generate a CSV.
</div>

```{r inititalize-data-structure}
  tournament = data.frame(Player = integer(), 
                             Name = character(), 
                             State = character(), 
                             Points = double(), 
                             Pre_Rtg = integer(),
                             Opp_Rtg = double(),
                             Games = integer(),
                             Opponents = character(),
                             stringsAsFactors = FALSE)
```

## Processing

<div id = "solution">
This is the meat of the work. Extensive use of R's chaining and nesting capabilities were done to produce a single statement to produce most variables. I created a function to "camel case" the players' names as per the project requirements.
</div>

```{r process-file}

# created function to convert names to camel case format
# preventing the need to loop through the segments of the name
camelCase = function(s){
  # converts the first letter of each word to upper case and the other characters to lower case
  return(str_c(toupper(str_sub(s,1,1)),tolower(str_sub(s,2,str_count(s))))) 
}

i <- 5 #start at the 5th line
while (i <= length( lines ) ){
   playerInfoRow1 <- str_trim(str_split(str_trim(lines[i]),"\\|")[[1]])
   playerInfoRow2 <- str_trim(str_split(str_trim(lines[i + 1]),"\\|")[[1]])
   playerNumber <- playerInfoRow1[1]
   playerName <- paste(unlist(camelCase(str_split(playerInfoRow1[2],"\\s")[[1]])), collapse = " ")
   playerState <- playerInfoRow2[1]
   points <- playerInfoRow1[3]
   playerRatings= str_trim(str_split(str_split(playerInfoRow2[2] ,":")[[1]][2],"->")[[1]])
   playerPreRating <- str_split(playerRatings[1],"P")[[1]][1]
   opponents= str_extract_all(str_sub(lines[i],-43),"[0-9]+")[[1]] #used as reference calc. average
   gamesPlayed = length(opponents)
   opponents=paste(unlist(opponents),collapse = "|")
   tournament[nrow(tournament) + 1,] = list(playerNumber, 
                                            playerName, 
                                            playerState,
                                            points, 
                                            playerPreRating, 
                                            0.0,
                                            gamesPlayed,
                                            opponents)
   
   # increment in blocks of 3 to prevent unnecessary processing of dashes
   i <- i + 3
}
```
## Preview Processed Data
<div id = "solution">
  We now all the necessary information to perform necessary calculations and analysis.
</div>

```{r preview-processed-data}
  print(subset(tournament, select = c("Name","State","Points","Pre_Rtg","Opponents")))
```

## Calculate Opponents Pre Chess Rating by Subsetting

<div id = "solution">
This was done by referencing the players opponents by creating a list and using the *%in%* operator as input to the mean function.
</div>

```{r opponents-pre-averages}
i <- 1
while(i <= nrow(tournament)){
  oppMean <- mean(as.integer(subset(tournament, 
                                   Player %in% as.integer(str_split(tournament$Opponents[i],
                                                                            "\\|")[[1]]),select=c("Pre_Rtg"))$Pre_Rtg))
  tournament$Opp_Rtg[i] <- oppMean
  i <- i + 1
}
```

## View Opponents Pre Chess Rating

```{r preview-averages}
  print(subset(tournament, select = c("Name","State","Points","Pre_Rtg","Opp_Rtg")))
```

## Write CSV File

<div id = "solution">
 Only the necessary variables are written to the CSV file.
</div>

```{r write-file}
write.csv(subset(tournament, select = c("Name","State","Points","Pre_Rtg","Opp_Rtg")), 
          file = "tournament.csv", quote = FALSE, row.names = FALSE)
```